\textbf{\Large Hybrid Client/Server Rendering with Automatic Proxy Model Generation}\\[0.2ex]

### 2017-10-04 tex/full_2017/paper2.tex
*fixed author *removed anonemous stuff

\author{\IEEEauthorblockN{~\\[-0.4ex]\large Jens Olav Nygaard\\[0.3ex]\normalsize}
\IEEEauthorblockA{SINTEF Digital\\
Applied Mathematics\\
Norway\\
Email: {\tt Jens.O.Nygaard@sintef.no}}
\and
\IEEEauthorblockN{~\\[-0.4ex]\large Jon Mikkelsen\\[0.3ex]\normalsize}
\IEEEauthorblockA{SINTEF Digital\\
Applied Mathematics\\
Norway\\
Email: {\tt Jon.Mikkelsen@sintef.no}}}

@@

\twocolumn[{%
\renewcommand\twocolumn[1][]{#1}%
\maketitle
\begin{center}
    \centering
		\begin{tabular}[b]{ccc}
    \includegraphics[width=0.32\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png} &
    \includegraphics[width=0.32\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png} &
    \includegraphics[width=0.32\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png} \\
		a) & b) & c)
		\end{tabular}
		\captionof{figure}{\label{fig:FRView} The oil reservoir viewer YYYY~\cite{cloudviz},
  (name withheld for the double-blind review process) showing one
  server-rendered and two client-rendered images of automatically generated and
  slightly rotated proxy models.  For~(b), parameters are chosen to produce the
  best possible image, and in~(c) we want to highlight artifacts and
  implementational details. See the text for further discussion.}
%    \protect \label{fig:wfedesign} }
%    \captionof{figure}{Test caption}
\end{center}%
}]

@@

Hand in hand with increasingly powerful rendering engines comes ever increasing
requirements on computational accuracy, power efficiency, data sizes, scaling
properties, etc. This is also reinforced by new cloud-based approaches and
wireless usage patterns. An effect of this is that interactivity still is a
difficult issue. We consider a client/server model for 3D rendering, addressing
latency, bandwidth and scaling problems in a novel way.

We introduce a {\em proxy model}, defined as a temporary model to be shown and
manipulated locally on a client while waiting for the appropriate image from a
connected server. Producing such proxy models can be difficult for many kinds of
3D data, like in our main case in which we have an oil reservoir viewer
(YYYY,~\cite{cloudviz}) that renders large {\em corner point grids}, together
with faults, oil wells, and more. See Figure~\ref{fig:FRView} above for an
example of both a full server-side rendering, and automatically generated proxy
models rendered in Google Chrome.
%
Our solution is to pass depth information from the server along with ordinary
rendered images. From this, a rudimentary 3D model is built, and with the RGB
image as a texture, this model can be manipulated and rendered on the client
while waiting for the next update from the server. If the client does not change
the position or orientation of the model too much, this proxy model rendering
integrates seamlessly with the slower stream of server-rendered frames.
%
Even if bandwidth and latency is not a problem, it may be desirable to let a
server of limited capacity serve many simultaneous users, hence limiting the
effective server time available for each one. Suitable scaling may still be
achieved using our solution. This is currently being commercialized in an
ongoing XXXX (name withheld for the double-blind review process) project.

@@

Our approach has some similarities to {\em image based rendering} (IBR)
techniques, with the difference that an important IBR problem would be the
reconstruction of a depth map from images, while we have access to the full
depth map from a rendering pass.  Another way to use depth maps similar to what
we do, is for ``immersive streaming'', see \eg,~\cite{ibr}, where focus is on
depth map compression, an issue that we also consider. Another work
focused on similar streaming and geometry compression,
is~\cite{teler}. Also,~\cite{220764} contains some of the same ideas as our
work, with respect to image-based rendering acceleration.
%
\textbf{A very early work aiming at the same kind of ``inter-frame rendering'', is
\cite{Mark:1997:PW:253284.253292}. Here, several frames are warped, and
subsequently combined, in order to avoid large unpainted areas caused by
occlusion. This corresponds to our use of several proxy models, each from a pair
of (rgb, depth)-images, the main difference being that they work on separate
pixels, while we use larger, textured ``splats''. Our method avoids their
slightly complicated meshing, discontinuity detection and final image
composition stage.}
%
Common to many IBR-algorithms is also that of stitching together 3D or 2.5D
point clouds. We could do this for our 2.5D maps fetched from the server, but it
is unclear if the benefit would outweigh the cost.
%
\textbf{[OLD: We are not aware of anybody else using depth maps exactly for our purpose, most
research seems to be focused on the retrieval of the depth map from images,
while we use an existing depth map to distill and render temporary geometry.]}
\textbf{[NEW: Our approach differs from many similar ones, in that we use existing depth maps 
to distill and render temporary geometry, rather than retrieveing the depth maps from images.]}
%
We observe that these 2.5D height maps are exactly what ``3D cameras''
(time-of-flight and other range image sensors) produce, but most authors
considering these are typically building more complex geometries before
visualization, see~\cite{IMM2009-05801}.
%
\textbf{Another approach is that of~\cite{altenhofen16rixels}. They use websockets
where we use the http protocol, a more significant difference is that we get
away with transferring a lot less data due to our adaptive compression ratio
depending on continuous bandwidth and latency measurements.
%
}
%
\textbf{In~\cite{CGF:CGF1871}, Pajak \etal considers remote rendering and
  streaming of frames rendered from a dynamical 3D model, we are quite agnostic
  to the source of imagery. Their setup requires more powerful clients than ours
  (OpenGL vs. WebGL), but they will also have higher fidelity.  }
%
\textbf{A special case is provided for in the rendering of stereo images. In
  this case, the 3D disparity is limited, while the rendering cost is doubled,
  since two views per frame is needed. In~\cite{DidykERMS2010}, this is
  exploited to make a solution tailored to such stereo synthesis; performance
  approaches that of rendering non-stereo, with a minimization of depth
  disparity artifacts. Occlusion and disocclusion holes are avoided by warping quads rather than
  pixel and filling in with previous images.}
%
% Er denne egentlig relevant?
%\textbf{Nalbach \etal,~\cite{Nalbach:2014:DSS:2556700.2556708}, ...
%}

@@

We have tested the algorithms discussed in the ZZZZ (name withheld for the
double-blind review process) framework, which is a programming framework for
setting up and managing client/server based interactive visualization
applications, see~\cite{tinia}. As client we use Google Chrome, code is written
in standard Javascript/WebGL.  The implementation in ZZZZ is invisible to the
application, so all existing ZZZZ-applications will have the feature
available. The algorithm is minimally intrusive in that only the depth buffer
will have to be added to the rendering output of the application. We have tested
several smaller test cases, but also on a larger oil reservoir viewer,
YYYY~\cite{cloudviz}.

YYYY utilizes several GLSL shaders to render reservoir cells, boundaries,
tubular wells etc., and visualizes a 3D model that is not trivial to reduce in
complexity. It is typically also very large, so rendering it on a thin client is
prohibitive. With the automatic proxy model, we obtain interactive frame rates
with limited connection from a lightweight client. For a comparison of a
server-rendered image and a client-rendered proxy model that is slightly rotated
on the client, see again Figure~\ref{fig:FRView}. In
Figure~\ref{fig:FRView}~(a), the full server-rendered image is shown, and in~(b)
and~(c), a slightly (about 10 degrees) rotated proxy model is rendered on the
client. In~(b), parameters are chosen for best possible results, while in~(c),
we want to highlight effects, so it uses a small number of non-overlapping large
splats. One can easily spot areas not well covered, and areas where the texture
coordinate transform $T$ does not produce optimal results.

>>>>

\author{\IEEEauthorblockN{~\\[-0.4ex]\large Jens Olav Nygaard\\[0.3ex]\normalsize}
\IEEEauthorblockA{SINTEF Digital\\
Applied Mathematics\\
Norway\\
Email: {\tt Jens.O.Nygaard@sintef.no}}
\and
\IEEEauthorblockN{~\\[-0.4ex]\large Jon Mikkelsen Hjelmervik\\[0.3ex]\normalsize}
\IEEEauthorblockA{SINTEF Digital\\
Applied Mathematics\\
Norway\\
Email: {\tt Jon.M.Hjelmervik@sintef.no}}}

@@

\twocolumn[{%
\renewcommand\twocolumn[1][]{#1}%
\maketitle
\begin{center}
    \centering
		\begin{tabular}[b]{ccc}
    \includegraphics[width=0.32\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png} &
    \includegraphics[width=0.32\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png} &
    \includegraphics[width=0.32\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png} \\
		a) & b) & c)
		\end{tabular}
		\captionof{figure}{\label{fig:FRView} Oil reservoir viewer showing one
  server-rendered and two client-rendered images of automatically generated and
  slightly rotated proxy models.  For~(b), parameters are chosen to produce the
  best possible image, and in~(c) we want to highlight artifacts and
  implementational details. See the text for further discussion.}
%    \protect \label{fig:wfedesign} }
%    \captionof{figure}{Test caption}
\end{center}%
}]

@@

Hand in hand with increasingly powerful rendering engines comes ever increasing
requirements on computational accuracy, power efficiency, data sizes, scaling
properties, etc. This is also reinforced by popular cloud-based approaches and
wireless usage patterns. An effect of this is that interactivity still is a
difficult issue. We consider a client/server model for 3D rendering, addressing
latency, bandwidth and scaling problems in a novel way.

We introduce a {\em proxy model}, defined as a temporary model to be shown and
manipulated locally on a client while waiting for the appropriate image from a
connected server. Producing such proxy models can be difficult for many kinds of
3D data, like in our main case in which we have an oil reservoir viewer
(YYYY,~\cite{cloudviz}) that renders large {\em corner point grids}, together
with faults, oil wells, and more. See Figure~\ref{fig:FRView} above for an
example of both a full server-side rendering, and automatically generated proxy
models rendered in Google Chrome.
%
Our solution is to pass depth information from the server along with ordinary
rendered images. From this, a rudimentary 3D model is built, and with the RGB
image as a texture, this model can be manipulated and rendered on the client
while waiting for the next update from the server. If the client does not change
the position or orientation of the model too much, this proxy model rendering
integrates seamlessly with the slower stream of server-rendered frames.
%
Even if bandwidth and latency is not a problem, it may be desirable to let a
server of limited capacity serve many simultaneous users, hence limiting the
effective server time available for each one. Suitable scaling may still be
achieved using our solution. This is currently being commercialized as a part of
the result of  a recently finished EU-project called CloudFlow.

@@

Our approach has some similarities to {\em image based rendering} (IBR)
techniques, with the difference that an important IBR problem would be the
reconstruction of a depth map from images, while we have access to the full
depth map from a rendering pass.  Another way to use depth maps similar to what
we do, is for ``immersive streaming'', see \eg,~\cite{ibr}, where focus is on
depth map compression, an issue that we also consider. Another work
focused on similar streaming and geometry compression,
is~\cite{teler}. Also,~\cite{220764} contains some of the same ideas as our
work, with respect to image-based rendering acceleration.
%
\textbf{A very early work aiming at the same kind of ``inter-frame rendering'', is
\cite{Mark:1997:PW:253284.253292}. Here, several frames are warped, and
subsequently combined, in order to avoid large unpainted areas caused by
occlusion. This corresponds to our use of several proxy models, each from a pair
of (rgb, depth)-images, the main difference being that they work on separate
pixels, while we use larger, textured ``splats''. Our method avoids their
slightly complicated meshing, discontinuity detection and final image
composition stage.}
%
Common to many IBR-algorithms is also that of stitching together 3D or 2.5D
point clouds. We could do this for our 2.5D maps fetched from the server, but it
is unclear if the benefit would outweigh the cost.
%
Our approach differs from many similar ones, in that we use existing depth maps 
to distill and render temporary geometry, rather than retrieveing the depth maps from images.
%
We observe that these 2.5D height maps are exactly what ``3D cameras''
(time-of-flight and other range image sensors) produce, but most authors
considering these are typically building more complex geometries before
visualization, see~\cite{IMM2009-05801}.
%
\textbf{Another approach is that of~\cite{altenhofen16rixels}. They use websockets
where we use the http protocol, a more significant difference is that we get
away with transferring a lot less data due to our adaptive compression ratio
depending on continuous bandwidth and latency measurements.
%
}
%
\textbf{In~\cite{CGF:CGF1871}, Pajak \etal considers remote rendering and
  streaming of frames rendered from a dynamical 3D model, we are quite agnostic
  to the source of imagery. Their setup requires more powerful clients than ours
  (OpenGL vs. WebGL), but they will also have higher fidelity.  }
%
\textbf{A special case is provided for in the rendering of stereo images. In
  this case, the 3D disparity is limited, while the rendering cost is doubled,
  since two views per frame is needed. In~\cite{DidykERMS2010}, this is
  exploited to make a solution tailored to such stereo synthesis; performance
  approaches that of rendering non-stereo, with a minimization of depth
  disparity artifacts. Occlusion and disocclusion holes are avoided by warping quads rather than
  pixel and filling in with previous images.}
%
% Er denne egentlig relevant?
%\textbf{Nalbach \etal,~\cite{Nalbach:2014:DSS:2556700.2556708}, ...
%}

@@

We have tested the algorithms proposed algorithms through the Tinia 
framework~\cite{tinia}, which is a programming framework for
setting up and managing client/server based interactive visualization
applications. As client we use Google Chrome, code is written
in standard Javascript/WebGL.  The auto-proxy implementation in is invisible to the
application, so all existing Tinia-applications will have the feature
available. The algorithm is minimally intrusive in that only the depth buffer
will have to be added to the rendering output of the application. We have tested
several smaller test cases, but also on a larger oil reservoir viewer.

The reservoir viewer utilizes several GLSL shaders to render reservoir cells, boundaries,
tubular wells etc., and visualizes a 3D model that is not trivial to reduce in
complexity. It is typically also very large, so rendering it on a thin client is
prohibitive. With the automatic proxy model, we obtain interactive frame rates
with limited connection from a lightweight client. For a comparison of a
server-rendered image and a client-rendered proxy model that is slightly rotated
on the client, see again Figure~\ref{fig:FRView}. In
Figure~\ref{fig:FRView}~(a), the full server-rendered image is shown, and in~(b)
and~(c), a slightly (about 10 degrees) rotated proxy model is rendered on the
client. In~(b), parameters are chosen for best possible results, while in~(c),
we want to highlight effects, so it uses a small number of non-overlapping large
splats. One can easily spot areas not well covered, and areas where the texture
coordinate transform $T$ does not produce optimal results.
